---
title: "ridge_regression"
author: "Lynn Chen"
date: "1/17/2022"
output: html_document
---

```{r}
library(tidyverse)
library(MASS)
library(performance) # vif
library(caret)
library(car)
library(Metrics)
library(corrplot)
library(glmnet)
library(rsample)

```


```{r}
le_df = 
  read_csv("./data/Dummy_Data.CSV") 
```

## Ridge Regression
```{r}
#define response variable
y = le_df$le

#define matrix of predictor variables
x = data.matrix(le_df[, c('status_Developed', 'adult_mort', 'infantdeaths', 'alcohol', 'percent_exp', 'hepatitis_b_high', 'measles', 'bmi', 'under_five', 'polio_high', 'total_exp', 'diphtheria_high', 'hiv_aids', 'gdp', 'population', 'thin_1_19', 'thin_5_9', 'HDI', 'schooling')])

#fit ridge regression model
ridge_model = glmnet(x, y, alpha = 0)

#view summary of model
summary(ridge_model)
dim(coef(ridge_model))
plot(ridge_model) 

ridge_model$lambda[50] 
coef(ridge_model)[,50] 
sqrt(sum(coef(ridge_model)[-1,50]^2)) # Calculate l2 norm
predict(ridge_model, s = 50, type = "coefficients")[1:20,]

```

## Identify lambda value 

Find lambda value that produces the lowest test mean squared error (MSE) by using k-fold cross-validation.

```{r}
#perform k-fold cross-validation to find optimal lambda value
cv_model = cv.glmnet(x, y, alpha = 0)

#find optimal lambda value that minimizes test MSE
best_lambda = cv_model$lambda.min
best_lambda

#produce plot of test MSE by lambda value
plot(cv_model) 
```

* The lambda value that minimizes the test MSE turns out to be 0.7175106.
* As λ increases, the flexibility of the ridge regression fit decreases, leading to decreased variance but increased bias. Big lambda -- underfitting

## Analyze Final Model

```{r}
#find coefficients of best model
best_model = glmnet(x, y, alpha = 0, lambda = best_lambda)
coef(best_model)

#produce Ridge trace plot
plot(ridge_model, xvar = "lambda")

#use fitted best model to make predictions
y_predicted = predict(best_model, s = best_lambda, newx = x)

#find SST and SSE
sst = sum((y - mean(y))^2)
sse = sum((y_predicted - y)^2)

#find R-Squared
rsq = 1 - sse/sst
rsq

```
* best model was able to explain 79.99% of the variation in the response values of the training data.

## Split into training set and a test set

```{r}
set.seed(1)

train = le_df %>%
  sample_frac(0.75)

test = le_df %>%
  setdiff(train)

x_train = model.matrix(le ~ status_Developed + adult_mort + infantdeaths + alcohol + percent_exp + hepatitis_b_high + measles + bmi + under_five + polio_high + total_exp + diphtheria_high + hiv_aids + gdp + population + thin_1_19 + thin_5_9 + HDI + schooling, train)[,-1]

x_test = model.matrix(le ~ status_Developed + adult_mort + infantdeaths + alcohol + percent_exp + hepatitis_b_high + measles + bmi + under_five + polio_high + total_exp + diphtheria_high + hiv_aids + gdp + population + thin_1_19 + thin_5_9 + HDI + schooling, test)[,-1]

y_train = train$le

y_test = test$le  

ridge_mod = glmnet(x_train, y_train, alpha=0, lambda = best_lambda, thresh = 1e-12)
ridge_pred = predict(ridge_mod, s = 4, newx = x_test)
mean((ridge_pred - y_test)^2)

ridge_pred = predict(ridge_mod, s = 0, newx = x_test, exact = F)
mean((ridge_pred - y_test)^2)

lm(le ~ status_Developed + adult_mort + infantdeaths + alcohol + percent_exp + hepatitis_b_high + measles + bmi + under_five + polio_high + total_exp + diphtheria_high + hiv_aids + gdp + population + thin_1_19 + thin_5_9 + HDI + schooling, data = train)
predict(ridge_mod, s = 0, exact = F, type="coefficients")[1:20,]
```

## RMSE

```{r}
#splitting data into testing and training sets
set.seed(123)
initial_split <- initial_split(le_df, prop = 0.75)
expectancy_train <- training(initial_split)
expectancy_test <- testing(initial_split)

train_control = trainControl(method="cv", number = 5)

model_ridge_cv = train(le ~ status_Developed + adult_mort + infantdeaths + alcohol + percent_exp + hepatitis_b_high + measles + bmi + under_five + polio_high + total_exp + diphtheria_high + hiv_aids + gdp + population + thin_1_19 + thin_5_9 + HDI + schooling, data = expectancy_train[,-1], trControl=train_control, method="glmnet", preProcess = c("center", "scale"),tuneGrid = expand.grid(alpha = 0, lambda = best_lambda))

ridge_pred = predict(model_ridge_cv, model.matrix(le ~ status_Developed + adult_mort + infantdeaths + alcohol + percent_exp + hepatitis_b_high + measles + bmi + under_five + polio_high + total_exp + diphtheria_high + hiv_aids + gdp + population + thin_1_19 + thin_5_9 + HDI + schooling, expectancy_test)[,-1])

ridge_rmse = rmse(ridge_pred,expectancy_test$le)

ridge_rmse

```

